---
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi Outcomes AugSynth Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
library(kableExtra)
```

# `augsynth`: Estimating multiple outcome effects

### The data
To demonstrate `augsynth` with multiple outcomes, we'll use data on the impact of personal income tax cuts in Kansas that comes with the `AugSynth` package. Our interest is in estimating the effect of income tax cuts on gross state product (GSP) per capita, employment, wages, and other macroeconomic indicators.

```{r load_data, results="hide", warning=F, message=F}
library(magrittr)
library(dplyr)
library(augsynth)
data(kansas)
```

The `kansas` dataset contains the GSP per capita (the outcome measure) `lngdpcapita` for all 50 states from the first quarter of 1990 to the first quarter of 2016.

To run `augsynth`, we need to include a treatment status column that indicates which region was treated and at what time. The table in `kansas` contains the column `treated` to denote this. In the original study, the second quarter of 2012 was the implementation of the tax cut in Kansas.

```{r treated_units}
kansas %>% select(year, qtr, year_qtr, state, treated, gdp, lngdpcapita) %>% filter(state == "Kansas" & year_qtr >= 2012 & year_qtr < 2013) 
```


### The single outcome case
We will begin by running the synthetic controls method and the augmented synthetic controls method on our outcome variables separately.
