---
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{MultiSynth Vignette}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
    title: "Comparing inference approaches for the synthetic control method with the `augsynth` package"
subtitle: Replicating Abadie et al. (2010)
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
    collapse = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = 'center',
    comment = "#>"
)

## Install Synth if not already installed
# install.packages("Synth")

library(kableExtra)
library(magrittr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(rlang)

library(foreign)
library(graphics)
library(Synth)
library(augsynth)
data(kansas)

## For the Abadie plots
replication_theme <- theme_bw(base_family = "Times New Roman") + 
    theme(panel.grid = element_blank(), 
          legend.background = element_rect(color = 'black'),
          legend.spacing.y = unit(0.0, 'pt'),
          legend.box.margin = margin(t = 5, r = 5, unit = 'pt'),
          legend.position = c(0.81, 0.9))

set.seed(1234)
```


There are many competing types of inference in the Synthetic Control world, which rely on a variety of different sets of assumptions that researchers can get quite agitated about.
In this post, we are going to re-analyze the famous "Tobacco example" originally laid out in [Abadie et al. (2010)](https://web.stanford.edu/~jhain/Paper/JASA2010.pdf), using some cool new features we recently added to the `augsynth` package.

The synthetic control approach has two major steps.
The first step is to generate the synthetic control.
Intuitively, this consists of using some sort of algorithm to reweight a set of possible "donor" units so that their weighted average "looks like" the target treated unit, usually with respect to the treated unit's pre-treatment outcome trajectory.
There are an ever-increasing variety of ways one might do this, which we will discuss a bit more below.
The second step is to figure out how to conduct inference: we have deliberately manufactured a "synthetic" unit to look like our treated unit, and now we want to know if the difference in their behaviors post-treatment could be due to random chance, or something structural like a treatment effect.

In terms of R packages to do all of this for you, we note two contenders.
The first is the `Synth` package, which has the algoirthm and approach outlined in the seminal Abadie et al. (2010).
The second is the newer `augsynth` package, that offers a variety of different methods for building a synthetic comparison, and in fact _does not_ offer the exact same option used in `Synth`.
Nevertheless, we can fairly closely mimic the spirit of the "Original Synth" with `augsynth`, which gets us fairly close to the same results, as measured by Figures 2--7 in the 2010 paper, even though we are using a slightly different algorithm.

So what is the difference?
In short, unlike the `augsynth` package, the `Synth` package used in Abadie et al.'s analysis balances lagged outcomes *and* auxiliary covariates in selecting synthetic weights.
The Rmd version of this notebook includes `Synth` code to generate Abadie et al.'s results with a high degree of fidelity (with the resulting `Synth`-based plots shown adjacent to analogous output from `augsynth` models).


```{r import tobacco data, include=FALSE}
# Get working data CA and 38 control states 
Wk.data <- read.dta(here::here("./tobacco_replication/data/smoking_wkdata_39.dta"))

# load in "Rest of US" data (computed from 50 states) for figure 1
RestUS.50 <- read.dta(here::here("./tobacco_replication/data/restofus_50.dta"))

# this data is used for table 1, where we average the 38 control states
RestUS.38 <- read.dta(here::here("./tobacco_replication/data/restofus_38.dta"))
```

```{r Synth dataprep and model, echo=FALSE, cache=TRUE, include=FALSE}
### Run Baseline Model

dataprep.out <- dataprep(
    foo = Wk.data,
    predictors = c("retprice", "xxincome", "K1_r_15_24"),
    predictors.op = c("mean"),
    dependent = c('cigsalepcap'),
    unit.variable = c('index'),
    time.variable = c('year'),
    special.predictors = list(list("xxbeer", 1984:1988, c("mean")),
                              list("cigsalepcap", 1988, c("mean")),
                              list("cigsalepcap", 1980, c("mean")),
                              list("cigsalepcap", 1975, c("mean"))
    ),
    treatment.identifier = 3,
    controls.identifier = c(1:39)[-3],
    time.predictors.prior = 1980:1988,
    time.optimize.ssr = 1970:1988,
    unit.names.variable = c('name'),
    time.plot = 1970:2005
)

### run synth

synth.out <- synth(
    dataprep.out,
    Pop.size = 300,
    Max.generations = 1,
    Unif.seed = 356987,
    Int.seed = 627478,
    Optim.method = "BFGS",
    L.ipop = 0,
    Maxiter.ipop = 1e10,
    Margin.ipop = 0.05,
    Sigf.ipop = 8,
    genoud = TRUE
)
```

```{r Abadie Figure 2, echo=FALSE, include=FALSE, fig.height=4, fig.width=4.6}
f2_plot_df <- cbind(rownames(dataprep.out$Y1plot),
                    dataprep.out$Y1plot,
                    dataprep.out$Y0plot %*% synth.out$solution.w) %>%
    as_tibble() %>%
    mutate_all(as.numeric)

colnames(f2_plot_df) <- c('year', 'California', 'Synthetic California')

f2_plot_df <- f2_plot_df %>%
    reshape2::melt(id.vars = 'year',
                   variable.name = 'treat_group',
                   value.name = 'per_cap_cigs')

path_plot <- ggplot(f2_plot_df, aes(x = year, y = per_cap_cigs, linetype = treat_group)) +
    geom_line() +
    geom_vline(linetype = 'dotted', xintercept = 1988) +
    scale_linetype_manual(values = c('solid', 'dashed'),
                          breaks = c("California", "Synthetic California")) +
    scale_y_continuous(breaks = seq(0, 141, 20), limits = c(0, 140)) +
    scale_x_continuous(breaks = seq(1970, 2005, 5), limits = c(1970, 2005)) +
    labs(linetype = NULL,
         title = 'Abadie et al. (2010), Figure 2',
         y = 'per-capita cigarette sales (in packs)') +
    annotate('text', y = 41, x = 1987.5, label = 'Passage of Proposition 99 \u2192',
             hjust = 1, color = 'black', size = 3, family = "Times New Roman") +
    annotate('text', y = 0, x = 1970, label = 'ND replication: Synth package',
             hjust = 0, color = 'darkseagreen', size = 3) +
    replication_theme

path_plot
```

```{r Abadie Figure 3, echo=FALSE, include=FALSE, fig.height=4, fig.width=4.6}
gap <- dataprep.out$Y1plot - dataprep.out$Y0plot %*% synth.out$solution.w
year <- dataprep.out$tag$time.plot

plot_df <- cbind(gap, year) %>% as_tibble()
colnames(plot_df) <- c('gap', 'year')

gap_plot <- ggplot(plot_df, aes(x = year, y = gap)) +
    geom_line() +
    geom_vline(linetype = 'dotted', xintercept = 1988) +
    geom_hline(linetype = 'dashed', yintercept = 0) +
    scale_y_continuous(breaks = seq(-30, 30, 10), limits = c(-30, 30)) +
    scale_x_continuous(breaks = seq(1970, 2005, 5), limits = c(1970, 2005)) +
    labs(linetype = NULL,
         title = 'Abadie et al. (2010), Figure 3',
         y = 'gap in per-capita cigarette sales (in packs)') +
    annotate('text', y = -25, x = 1987.5, label = 'Passage of Proposition 99 \u2192',
             hjust = 1, color = 'black', size = 3, family = "Times New Roman") +
    annotate('text', y = -30, x = 1970, label = 'ND replication: Synth package',
             hjust = 0, color = 'darkseagreen', size = 3) +
    replication_theme

gap_plot
```

### Model setup

We can now execute an `augsynth` model with or without inference. Inference options include those already included in the package (`conformal`, `jackknife`, `jackknife+`), but also `permutation` and `permutation_rstat` (RMSPE-adjusted permutation inference), which are calculated using placebo tests. 

All inference results, similar to those previously returned by `summmary.augsynth()`, are stored in `augsynth$results` by default when the `inf_type` argument is specified. 

```{r reformat tobacco data using tidyverse, echo=FALSE}
tobacco <- Wk.data %>%
    mutate(treated = ifelse((name == 'California') &
                                (year > 1988), 1, 0),
           state = name, # rename something meaningful
           state_index = index) %>%
    select(cigsalepcap, treated, state, year, retprice, xxincome, K1_r_15_24, xxbeer)

tobacco_70 <- tobacco %>% filter(year >= 1970) # subset data to 1970 forward to avoid issues due to missingess of outcomes
```

Note that if we use `progfunc = "none"` and `scm = TRUE` then the results deviate more strongly from the tobacco paper, but if we allow augsynth to do its default bias removal, we more closely replicate the paper:

```{r run tobacco model using augsynth}
# Basic SCM with covariates
syn <- augsynth(form = cigsalepcap ~ treated | retprice + xxincome + K1_r_15_24,
                unit = state,
                time = year,
                data = tobacco_70
                #progfunc = "none", scm=TRUE
)
```

### Summmarizing results

Calling `summary()` on an `augsynth` object can now accommodate an expanded range of inference types. 

**Question:** Currently, jackknife inference returns standard errors but not confidence intervals for the ATT estimate. Why is this?

```{r plot augsynth kansas models, echo=FALSE, cache=TRUE, fig.width=8, fig.height=4.5}
all_results <- bind_rows(Conformal = summary(syn, inf_type = 'conformal')$att, 
                         Jackknife = summary(syn, inf_type = 'jackknife')$att, 
                         `Jackknife+` = summary(syn, inf_type = 'jackknife+')$att, 
                         Permutation = summary(syn, inf_type = 'permutation')$att, 
                         `Permutation (rstat)` = summary(syn, inf_type = 'permutation_rstat')$att, 
                         .id = 'name') %>% 
    mutate(name = factor(name, 
                         levels = c('Conformal', 'Jackknife', 'Jackknife+', "Permutation", "Permutation (rstat)"), 
                         ordered = T))

ggplot(all_results, aes(x = Time, y = Estimate, color = name)) + 
    geom_hline(yintercept = 0, linetype = 'solid') + 
    geom_vline(xintercept = syn$t_int, linetype = 'dashed') + 
    geom_line() +
    geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = name), alpha = 0.2, size = 0.1) + 
    facet_wrap(. ~ name) + 
    scale_x_continuous(breaks = seq(1970, 2010, 5)) + 
    labs(x = 'Year', y = 'Gap in per-capita cigarette sales\n(in packs)', 
         caption = 'ND replication: augsynth package') +
    scale_color_manual(values = c('red', 'blue', 'darkgreen', 'purple', 'orange'), 
                       guide = 'none', aesthetics = c('color', 'fill')) +
    theme_bw() +
    theme(plot.caption = element_text(hjust = 0, color = 'darkseagreen', size = 9))
```


### Permutation inference

Adding permutation inference (via `inf_type = "permutation"` or `inf_type = "permutation_rstat"`) will return information about the 95% confidence interval, as well as other data generated from the raw placebo tests. Placebo results can be gotten via the `placebo_distribution()` method:

```{r show_perm_inference}
syn_rstat = summary( syn, inf_type = "permutation_rstat" )
placebo_distribution( syn_rstat )
```


### Plotting `augsynth` models

Generally one should plot the summarized `augsynth` model, as the summary has the inferenence. 
The proposed extensions to plotting capabilities include the possibility to customize the inferential features of plots and to easily generate plots showing the permuted ATTs from placebo tests, when appropriate.

If `plot()` is called on a summarized model object without additional arguments, it will return an ATT estimate with a 95% confidence interval calculated using conformal inference.
You can directly call plot on the original, pre-summarized, result, but this is not recommended as it will then have to calculate the inference on the fly:

```{r fig.width=4.5, fig.height=3.25}
plot(syn) + labs(title = 'Conformal inference (default)')
```

Setting `plot_type = "placebo"` on a summary with inference types of `permutation` or `permutation_rstat` will return a plot showing each placebo test. Otherwise, for other inference types, it will return only the ATT estimate. 

```{r fig.width=4.5, fig.height=4}
plot(syn_rstat, 
     plot_type = 'placebo',
     inf_type = 'permutation',
) + ggtitle("Replication of Abadie et al. (2010), Figure 4")
```


```{r  fig.width=4.5, fig.height=3.25}
p <- plot(syn_rstat, plot_type = 'estimate only') + ggtitle("Replication of Abadie et al. (2010), Figure 3")
```

```{r Format figures 3 comparison, echo=FALSE, fig.show="hold", fig.align='default', fig.height=4, fig.width=4.6}
gap_plot
p
```



If an outcomes plot is desired, the user can call `plot(plot_type = 'outcomes')` on an `augsynth` object. 

```{r  fig.width=4.5, fig.height=4}
p <- plot(syn_rstat, plot_type = 'outcomes') + 
    ggtitle("Replication of Abadie et al. (2010), Figure 2")
```

```{r Format figures 2 comparison, echo=FALSE, fig.show="hold", fig.align='default', fig.height=4, fig.width=4.6}
path_plot
p 
```

Alternatively, `plot(augsynth, plot_type = 'outcomes raw average')` will return both the raw average of donor units and the synthetic counterfactual.

```{r Outcomes plot with average, echo=TRUE, fig.show="hold", fig.align='default', fig.height=4, fig.width=4.6}
plot(syn_rstat, plot_type="outcomes raw average")
```

### Examining donor units and RMSPE

The `donor_table()` function will return a summary dataframe with the RMSPE and synthetic weight for each of the donor units.

```{r}
donor_table(syn_rstat) %>%
    arrange( -abs(weight) )
```


### Limiting the donor pool

The augsynth models can be re-fitted after excluding certain donor units via the `update_augsynth()` method.
If `drop` is passed to `update_augsynth()` as a numeric, then donor units will be excluded if their RMSPE is greater than a certain multiple of the treated unit's pre-treatment RMSPE. For example, setting `drop = 2` will exclude any units with a pre-treatment RMSPE twice as large as California's.
Donor units can also be specified using their unit identifiers (ex. "New Hampshire").

Updated augsynth models inherit the inference type, formula, and data structure of the original model.

```{r Figures 6 and 7, echo=TRUE, eval=TRUE, fig.show="hold", fig.align='default', fig.height=4.5, fig.width=4.6}
update_augsynth(syn) %>% # drops units with >20x treated RMSPE by default
    summary( inf_type = 'permutation' ) %>%
    plot(plot_type = 'placebo' ) + 
    ggtitle("Replication of Abadie et al. (2010), Figure 5") + ylim(-51, 91) + 
    annotate('text', y = -48, x = 1970, label = "Removes donors with 20x California's  pre-treatment RMSPE",
             hjust = 0, color = 'darkseagreen', size = 3) 
update_augsynth(syn, drop = 5) %>% 
    summary( inf_type = 'permutation' ) %>%
    plot(plot_type = 'placebo') + 
    ggtitle("Replication of Abadie et al. (2010), Figure 6") + ylim(-51, 91) + 
    annotate('text', y = -48, x = 1970, label = "Removes donors with 5x California's pre-treatment RMSPE",
             hjust = 0, color = 'darkseagreen', size = 3) 
update_augsynth(syn, drop = 2) %>% 
    summary( inf_type = 'permutation' ) %>%
    plot(plot_type = 'placebo') + 
    ggtitle("Replication of Abadie et al. (2010), Figure 7") + ylim(-51, 91) + 
    annotate('text', y = -48, x = 1970, label = "Removes donors with 2x California's pre-treatment  RMSPE",
             hjust = 0, color = 'darkseagreen', size = 3) 
```

The names of units to be excluded can be passed in directly as characters using `drop`. 

```{r permutation without states, echo=TRUE, eval=TRUE, fig.show="hold", fig.align='default', fig.height=4.5, fig.width=4.6}
drop_states <- c("Iowa", "Arizona", "Alabama", "Illinois", "Indiana", "Idaho", "Connecticut", 
                 "New Mexico", "Texas", "Utah", "North Dakota", "South Dakota", "Vermont", 
                 "Wisconsin", "West Virginia", "Wyoming", "Tennessee", "Pennsylvania")

update_augsynth(syn, drop = drop_states) %>% 
    summary( inf_type = 'permutation' ) %>%
    plot(plot_type = 'placebo') 
```

