---
title: "Permutation inference for the `augsynth` package"
subtitle: "Replicating Abadie et al. (2010)"
author: "Nora Delaney"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center',
  comment = "#>"
  )

## Install Synth if not already installed
# install.packages("Synth")

library(kableExtra)
library(magrittr)
library(dplyr)
library(ggplot2)
library(haven)
library(reshape2)
library(tidyr)
library(rlang)
 
library(foreign)
library(graphics)
library(Synth)
library(augsynth)
data(kansas)

## For the Abadie plots
replication_theme <- theme_bw(base_family = "Times New Roman") + 
  theme(panel.grid = element_blank(), 
        legend.background = element_rect(color = 'black'),
        legend.spacing.y = unit(0.0, 'pt'),
        legend.box.margin = margin(t = 5, r = 5, unit = 'pt'),
        legend.position = c(0.81, 0.9))

set.seed(1234)
```

# Project Overview

This notebook demonstrates the functionality of several extensions to the `augsynth` package, using the data and analytic context laid out in [Abadie et al. (2010)](https://web.stanford.edu/~jhain/Paper/JASA2010.pdf). The package extensions include the option to use permutation-based with models generated using `augsynth()`. In addition, the changes expand the plotting options available for `augsynth()` objects.

## Summary of changes

This project currently lives on the [n-delaney/augsynth_permutation](https://github.com/n-delaney/augsynth_permutation) GitHub repository. All changes should (ideally) be backwards compatible, although they are only currently implemented for cases where `single_augsynth()` is executed. The proposed extensions to the `augsynth` package do the following:

* Add the option to include inference in `augsynth` objects (in addition to summary objects). Inference details are stored in  `augsynth$results`.
* Extend the inference options available for `augsynth` objects with a single treated unit. Additional options now include placebo-based inference through the `inf_type = permutation` and `inf_type = permutation_rstat` (RMSPE-adjusted placebo tests). 
* Modify  `summary.augsynth()` 
  * If an `augsynth` does not already have inference results, `summary.augsynth()` will calculate them. 
  * It's also possible to use `summary.augsynth()` to re-do inference on an existing model with different characteristics (for example, switching from the default of "conformal" to "permutation" inference).
  * Moves inference calculations from `summary.augsynth()` to `add_inference()`, a function that handles inference calculations, whether executed as part of  `augsynth()` or  `summary.augsynth()`
* Add flexibility to `plot.augsynth()` 
  * Allow users to choose from or update inference types for a given `augsynth` model (ex. comparing plots using conformal and placebo inference).
  * Allow users to select from a number of stylized plots using the `plot_type` argument (the default is the ATT estimate and 95% confidence intervavl):
    * "estimate" — The ATT and 95% confidence interval
    * "estimate only" — The ATT without a confidence interval
    * "outcomes" — The level of the outcome variable for the treated and synthetic control units.
    * "outcomes raw average" — The level of the outcome variable for the treated and synthetic control units, along with the raw average of the donor units.
    * "placebo" — The ATTs resulting from placebo tests on the donor units.
  * `plot.summary.augsynth()` should still be backwards compatible but no longer handles most of the plotting. 
* Add `augsynth_gap_plot()`, which shows the outcome trajectory of the treated unit, its synthetic counterfactual, and (optionally) the simple average of donor units.

# Replication

This vignette replicates the analysis and figures in Abadie et al. (2010) using the `augsynth` package. Whereas the original paper relies on the `Synth` package, the proposed extensions to the `augsynth` package are intended to streamline permutation-based inference and associated visualizations on behalf of the user. The following vignette shows how to use the `augsynth` package, including extended inference types and plotting functionality, to generate results similar (but not entirely identical) to Figures 2-7 in the original paper.

While the general intuition of Abadie et al.'s analysis can be replicated using the `augsynth` package, it is worth noting that covariates and model parameters are handled somewhat differently than in the original analysis. In particular, unlike the `augsynth` package, the `Synth` package used in Abadie et al.'s analysis balances lagged outcomes *and* auxiliary covariates in selecting synthetic weights. The Rmd version of this notebook includes `Synth` code to generate Abadie et al.'s results with a high degree of fidelity (with the resulting `Synth`-based plots shown adjacent to analogous output from `augsynth` models).

```{r import tobacco data, include=FALSE}
# Get working data CA and 38 control states 
Wk.data <- read.dta(here::here("./tobacco_replication/data/smoking_wkdata_39.dta"))

# load in "Rest of US" data (computed from 50 states) for figure 1
RestUS.50 <- read.dta(here::here("./tobacco_replication/data/restofus_50.dta"))

# this data is used for table 1, where we average the 38 control states
RestUS.38 <- read.dta(here::here("./tobacco_replication/data/restofus_38.dta"))
```

```{r Synth dataprep and model, echo=FALSE, cache=TRUE, include=FALSE}
### Run Baseline Model

dataprep.out <- dataprep(
  foo = Wk.data,
  predictors = c("retprice", "xxincome", "K1_r_15_24"),
  predictors.op = c("mean"),
  dependent = c('cigsalepcap'),
  unit.variable = c('index'),
  time.variable = c('year'),
  special.predictors = list(list("xxbeer", 1984:1988, c("mean")),
                            list("cigsalepcap", 1988, c("mean")),
                            list("cigsalepcap", 1980, c("mean")),
                            list("cigsalepcap", 1975, c("mean"))
                            ),
  treatment.identifier = 3,
  controls.identifier = c(1:39)[-3],
  time.predictors.prior = 1980:1988,
  time.optimize.ssr = 1970:1988,
  unit.names.variable = c('name'),
  time.plot = 1970:2005
  )

### run synth

synth.out <- synth(
                   dataprep.out,
                   Pop.size = 300,
                   Max.generations = 1,
                   Unif.seed = 356987,
                   Int.seed = 627478,
                   Optim.method = "BFGS",
                   L.ipop = 0,
                   Maxiter.ipop = 1e10,
                   Margin.ipop = 0.05,
                   Sigf.ipop = 8,
                   genoud = T
                   )
```

```{r Abadie Figure 2, echo=FALSE, include=FALSE, fig.height=4, fig.width=4.6}
f2_plot_df <- cbind(rownames(dataprep.out$Y1plot),
                 dataprep.out$Y1plot,
                 dataprep.out$Y0plot %*% synth.out$solution.w) %>%
  as_tibble() %>%
  mutate_all(as.numeric)

colnames(f2_plot_df) <- c('year', 'California', 'Synthetic California')

f2_plot_df <- f2_plot_df %>%
  reshape2::melt(id.vars = 'year',
                 variable.name = 'treat_group',
                 value.name = 'per_cap_cigs')

path_plot <- ggplot(f2_plot_df, aes(x = year, y = per_cap_cigs, linetype = treat_group)) +
  geom_line() +
  geom_vline(linetype = 'dotted', xintercept = 1988) +
  scale_linetype_manual(values = c('solid', 'dashed'),
                        breaks = c("California", "Synthetic California")) +
  scale_y_continuous(breaks = seq(0, 141, 20), limits = c(0, 140)) +
  scale_x_continuous(breaks = seq(1970, 2005, 5), limits = c(1970, 2005)) +
  labs(linetype = NULL,
       title = 'Abadie et al. (2010), Figure 2',
       y = 'per-capita cigarette sales (in packs)') +
  annotate('text', y = 41, x = 1987.5, label = 'Passage of Proposition 99 \u2192',
           hjust = 1, color = 'black', size = 3, family = "Times New Roman") +
  annotate('text', y = 0, x = 1970, label = 'ND replication: Synth package',
           hjust = 0, color = 'darkseagreen', size = 3) +
  replication_theme

path_plot

ggsave(filename = here::here('./tobacco_replication/plots/synth_figure2.png'),
       height = 4, width = 4.6, units = 'in', dpi = 100)
```

```{r Abadie Figure 3, echo=FALSE, include=FALSE, fig.height=4, fig.width=4.6}
gap <- dataprep.out$Y1plot - dataprep.out$Y0plot %*% synth.out$solution.w
year <- dataprep.out$tag$time.plot

plot_df <- cbind(gap, year) %>% as_tibble()
colnames(plot_df) <- c('gap', 'year')

gap_plot <- ggplot(plot_df, aes(x = year, y = gap)) +
  geom_line() +
  geom_vline(linetype = 'dotted', xintercept = 1988) +
  geom_hline(linetype = 'dashed', yintercept = 0) +
  scale_y_continuous(breaks = seq(-30, 30, 10), limits = c(-30, 30)) +
  scale_x_continuous(breaks = seq(1970, 2005, 5), limits = c(1970, 2005)) +
  labs(linetype = NULL,
       title = 'Abadie et al. (2010), Figure 3',
       y = 'gap in per-capita cigarette sales (in packs)') +
  annotate('text', y = -25, x = 1987.5, label = 'Passage of Proposition 99 \u2192',
           hjust = 1, color = 'black', size = 3, family = "Times New Roman") +
  annotate('text', y = -30, x = 1970, label = 'ND replication: Synth package',
           hjust = 0, color = 'darkseagreen', size = 3) +
  replication_theme

gap_plot

ggsave(filename = here::here('./tobacco_replication/plots/synth_figure3.png'),
       height = 4, width = 4.6, units = 'in', dpi = 100)
```

### Model setup

We can now execute an `augsynth` model with or without inference. Inference options include those already included in the package (`conformal`, `jackknife`, `jackknife+`), but also `permutation` and `permutation_rstat` (RMSPE-adjusted permutation inference), which are calculated using placebo tests. 

All inference results (similar to those previously returned by `summmary.augsynth()` are stored in `augsynth$results` by default when `augsynth(..., inf = TRUE)`. 

```{r reformat tobacco data using tidyverse, echo=FALSE}
tobacco <- Wk.data %>%
  mutate(treated = ifelse((name == 'California') &
                            (year > 1988), 1, 0),
         state = name, # rename something meaningful
         state_index = index) %>%
  select(cigsalepcap, treated, state, year, retprice, xxincome, K1_r_15_24, xxbeer)

tobacco_70 <- tobacco %>% filter(year >= 1970) # subset data to 1970 forward to avoid issues due to missingess of outcomes
```

```{r run tobacco model using augsynth, cache=TRUE}
# Basic SCM with covariates
syn <- augsynth(form = cigsalepcap ~ treated | retprice + xxincome + K1_r_15_24,
                 unit = state,
                 time = year,
                 data = tobacco_70)

# Basic SCM with inference added during model execution
syn_rstat <- augsynth(form = cigsalepcap ~ treated | retprice + xxincome + K1_r_15_24,
                   unit = state,
                   time = year,
                   data = tobacco_70,
                   inf_type = 'permutation_rstat' # "conformal", "jackknife", "jackknife+", "permutation"
                   )

```

### Summmarizing results

Calling `summary()` on an `augsynth` object can now accommodate an expanded range of inference types. 

**Question:** Currently, jackknife inference returns standard errors but not confidence intervals for the ATT estimate. Why is this?

```{r plot augsynth kansas models, echo=FALSE, cache=TRUE, fig.width=8, fig.height=4.5}
all_results <- bind_rows(Conformal = summary(syn, inf_type = 'conformal')$att, 
                         Jackknife = summary(syn, inf_type = 'jackknife')$att, 
                         `Jackknife+` = summary(syn, inf_type = 'jackknife+')$att, 
                         Permutation = summary(syn, inf_type = 'permutation')$att, 
                         `Permutation (rstat)` = summary(syn, inf_type = 'permutation_rstat')$att, 
                         .id = 'name') %>% 
  mutate(name = factor(name, 
                       levels = c('Conformal', 'Jackknife', 'Jackknife+', "Permutation", "Permutation (rstat)"), 
                       ordered = T))

ggplot(all_results, aes(x = Time, y = Estimate, color = name)) + 
  geom_hline(yintercept = 0, linetype = 'solid') + 
  geom_vline(xintercept = syn$t_int, linetype = 'dashed') + 
  geom_line() +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound, fill = name), alpha = 0.2, size = 0.1) + 
  facet_wrap(. ~ name) + 
  scale_x_continuous(breaks = seq(1970, 2010, 5)) + 
  labs(x = 'Year', y = 'Gap in per-capita cigarette sales\n(in packs)', 
       caption = 'ND replication: augsynth package') +
  scale_color_manual(values = c('red', 'blue', 'darkgreen', 'purple', 'orange'), 
                     guide = 'none', aesthetics = c('color', 'fill')) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0, color = 'darkseagreen', size = 9))
```


If an `augsynth` already has inference details stored in `augsynth$results` and an inference type is not specified, the summary method will return the existing inference. Otherwise, the summary will default to using conformal inference (`inf_type = 'conformal'`)

```{r}
summary(syn_rstat)$inf_type
summary(syn)$inf_type
summary(syn, inf_type = 'jackknife')$inf_type
```

In addition to using `summary()`, a user can update or replace the inference type of an augsynth model using `add_inference()`.

```{r}
syn_new <- add_inference(syn, inf_type = 'jackknife')
syn_new$results %>% attributes()
```

### Permutation inference

Adding permutation inference (via `inf_type = "permutation"` or `inf_type = "permutation_rstat"`) will return information about the 95% confidence interval, as well as the raw placebo tests. Placebo results are stored in `syn_rstat$results$permutations`

```{r}
syn_rstat$results$permutations$placebo_dist
```


### Plotting `augsynth` models

The proposed extensions to plotting capabilities include the possibility to customize the inferential features of plots and to easily generate plots showing the ATTs from placebo tests, when appropriate. Like the updates to `summary.augsynth()`, changes to `plot.augsynth()` use the inferential features of the provided `augsynth` object, or otherwise allow users to specify the desired inferential feature. 

If `plot()` is called on a model object, it will return an ATT estimate with a 95% confidence interval calculated using conformal inference.

```{r fig.width=4.5, fig.height=3.25}
plot(syn) + labs(title = 'Conformal inference (default)')
```

Setting `plot_type = "placebo"` with inference types of `permutation` or `permutation_rstat` will return a plot showing each placebo test. Otherwise, for other inference types, it will return only the ATT estimate. 

```{r fig.width=4.5, fig.height=4}
plot(syn, 
     plot_type = 'placebo',
     inf_type = 'permutation',
     ) + ggtitle("Replication of Abadie et al. (2010), Figure 4")
```


```{r  fig.width=4.5, fig.height=3.25}
p <- plot(syn, plot_type = 'estimate only') + ggtitle("Replication of Abadie et al. (2010), Figure 3")
```

```{r Format figures 3 comparison, echo=FALSE, fig.show="hold", fig.align='default', fig.height=4, fig.width=4.6}
gap_plot
p
```



If an outcomes plot is desired, the user can call `plot(plot_type = 'outcomes')` on an `augsynth` object. The same result can be achieved using the `augsynth_outcomes_plot()` function.

```{r  fig.width=4.5, fig.height=4}
p <- plot(syn_rstat, plot_type = 'outcomes') + 
  ggtitle("Replication of Abadie et al. (2010), Figure 2")
```

```{r Format figures 2 comparison, echo=FALSE, fig.show="hold", fig.align='default', fig.height=4, fig.width=4.6}
path_plot
p 
```

By default, `augsynth_outcomes_plot()` will return both the raw average of donor units and the synthetic counterfactual.

```{r Gap plot with average, echo=TRUE, fig.show="hold", fig.align='default', fig.height=4, fig.width=4.6}
augsynth_outcomes_plot(syn_rstat)
```

### Examining donor units and RMSPE

The `donor_table()` function will return a summary dataframe with the RMSPE and synthetic weight for each of the donor units.

```{r}
donor_table(syn)
```


In addition, augsynth models can be customized to exclude certain donor units. If `drop` is passed to `update_augsynth()` as a numeric, then donor units will be excluded if their RMSPE is greater than a certain multiple of the treated unit's pre-treatment RMSPE. For example, setting `drop = 2` will exclude any units with a pre-treatment RMSPE twice as large as California's. Otherwise, donor units can be specified using their unit identifiers (ex. "New Hampshire").

Updated augsynth models will inherit the inference type, formula, and data structure of the original model.

```{r Figures 6 and 7, echo=TRUE, eval=TRUE, fig.show="hold", fig.align='default', fig.height=4.5, fig.width=4.6}
update_augsynth(syn) %>% # drops units with >20x treated RMSPE by default
  plot(plot_type = 'placebo', inf_type = 'permutation') + 
  ggtitle("Replication of Abadie et al. (2010), Figure 5") + ylim(-51, 91) + 
  annotate('text', y = -48, x = 1970, label = "Removes donors with 20x California's  pre-treatment RMSPE",
           hjust = 0, color = 'darkseagreen', size = 3) 
update_augsynth(syn, drop = 5) %>% 
  plot(plot_type = 'placebo', inf_type = 'permutation') + 
  ggtitle("Replication of Abadie et al. (2010), Figure 6") + ylim(-51, 91) + 
  annotate('text', y = -48, x = 1970, label = "Removes donors with 5x California's pre-treatment RMSPE",
         hjust = 0, color = 'darkseagreen', size = 3) 
update_augsynth(syn, drop = 2) %>% 
  plot(plot_type = 'placebo', inf_type = 'permutation') + 
  ggtitle("Replication of Abadie et al. (2010), Figure 7") + ylim(-51, 91) + 
  annotate('text', y = -48, x = 1970, label = "Removes donors with 2x California's pre-treatment  RMSPE",
           hjust = 0, color = 'darkseagreen', size = 3) 
```

The names of units to be excluded can be passed in directly as characters using `drop`. 

```{r permutation without states, echo=TRUE, eval=TRUE, fig.show="hold", fig.align='default', fig.height=4.5, fig.width=4.6}
drop_states <- c("Iowa", "Arizona", "Alabama", "Illinois", "Indiana", "Idaho", "Connecticut", 
                 "New Mexico", "Texas", "Utah", "North Dakota", "South Dakota", "Vermont", 
                 "Wisconsin", "West Virginia", "Wyoming", "Tennessee", "Pennsylvania")

update_augsynth(syn, drop = drop_states) %>% plot(plot_type = 'placebo', inf_type = 'permutation') 
```

